### 2019-10-28 _Abstract_ASHEcon_201910.Rmd
Initial analysis and abstract

---
title: "Decision Support and Health Insurance Choice"
author: Ian McCarthy & Evan Saltzman
date: October 2019
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
subparagraph: yes
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{titlesec}
   - \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
   - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
   - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
output: 
  pdf_document:
    keep_tex: TRUE
bibliography: 'D:/CloudStation/Professional/Bibliography/BibTeX_Library.bib'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stargazer, knitr, kableExtra,
               lfe)
```

```{r, include=FALSE}
load("data/R_workspace.Rdata")
```
### Background:
Health insurance markets are unique in many respects, not least of which is the increasing complexity of choosing an optimal health insurance plan. Such complexity has been well-documented in studies of health insurance choice in the Medicare Advantage market [@abaluck2011; @ketcham2012; @gruber2017]. One way to reduce the burden of this complexity is to provide professional decision support through private insurance agents or public assistance programs, both of which are available in the California health insurance exchanges created under the Affordable Care Act (ACA). In this paper, we examine the role of this decision assistance on health insurance plan choice and its implications for consumer welfare.

### Data:
We have data on `r format(tot.enroll, big.mark=",")` enrollee/year observations from the California health insurance exhanges from `r first.year` to `r last.year`. Our data include a variety of individual and household characteristics, plan choices and plan characteristics, and information on the type of decision assistance used by the enrollee (if any). From these data, we can identify each household's set of possible health insurance plans, and we employ premium and cost sharing subsidy formulas to calculate health insurance costs for each possible plan for each household.

### Methods:
We first estimate overall demand for health insurance using a nested logit discrete choice model, as in @saltzman2019. We then identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set. For example, Gold and Platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150% of the federal poverty level. We estimate the effect of decision assistance on dominated choices with a linear probability model, allowing for year and household fixed effects.

### Results:
Our nested logit models establish strong evidence that decision assistance matters for insurance choice. Turning specifically to dominated choices, our baseline results suggest that individuals with decision assistance are `r abs(round(reg2$coeff[1],4))*100` percentage points less likely to make a dominated choice. On a base of `r round(mean.dom.choice,3)*100`%, this reflects a `r abs(round(reg2$coeff[1]/mean.dom.choice,2))*100`% decrease in the probability of making a dominated choice.

We also find strong heterogeneities in the effects of different forms of decision assistance. In particular, we find that individuals using publicly provided decision assistance are `r abs(round(reg3$coeff[2],4))*100` percentage points less likely to select a dominated plan, while individuals using a private insurance agent are `r abs(round(reg3$coeff[1],4))*100` percentage points less likely to select a dominated plan.

### Discussion:
Our current results provide strong evidence that decision assistance has a significant and economically meaningful effect on health insurance plan choice. In terms of dominated choices, our early results suggest that this change in decision-making is welfare improving, and we continue to explore other measures of choice dominance to further confirm this result.

In future work, our identification strategy will exploit the Trump administration's removal of cost-sharing subisidies from the exchanges in 2018 and the subsequent response from insurers to increase premiums on their silver plans ("silver loading"). This response significantly changed the prevalence of dominated choices in each household's choice set, and in this way, the removal of the cost-sharing subsidies offers an exogenous change to the set of dominated plans available to each household.

We are also extending our analysis of the differential effects between public decision support versus assistance from private insurance agents/brokers. This analysis will determine whether insurance brokers are more likely to steer patients into plans offered by the sponsoring insurer.

<!-- Preliminary results further suggest strong heterogeneities across different types of decision assistance, with insurance-based decision support (i.e., insurance brokers) more likely to steer patients into plans offered by the sponsoring insurer. The steering effect appears sufficiently strong such that individuals using an insurance broker are no less likely to choose a dominated plan than individuals with no decision support. The reduction in prevalence of dominated choices is therefore driven entirely by publicly provided decision assistance. -->

# References