### 2021-03-22 finals/lunch-and-learn/lunch-learn-202103.Rmd
add presentation

---
title: "Insurance choice, decision assistance, and insurer steering"
subtitle: "Ian McCarthy & Evan Saltzman"
author: Emory University, Lunch and Learn
date: March 30, 2021 #"`r format(Sys.time(), '%d %B %Y')`"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
#knit: pagedown::chrome_print
---

<!-- Adjust some CSS code for font size, maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code, .remark-inline-code { 
    font-size: 20px;
}
</style>

<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(dplyr.summarise.inform = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, xaringanExtra)
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble()
xaringanExtra::use_panelset()
```

# Insurance Choice is Confusing

.center[
  ![](slate-2017.png)
]

.footnote[
  [Slate, 2017](https://slate.com/business/2017/08/americans-are-overpaying-for-insurance-because-obamacare-is-confusing.html)
]

---
count: false

# Insurance Choice is Confusing
.center[
  ![](health-affairs-2017b.png)
]

.footnote[
  [Health Affairs, January, 2017](https://www.healthaffairs.org/doi/10.1377/hlthaff.2016.0472)
]

---
count: false

# Insurance Choice is Confusing
.center[
  ![](health-affairs-2017.png)
]

.footnote[
  [Health Affairs, August, 2017](https://www.healthaffairs.org/doi/full/10.1377/hlthaff.2017.0085)
]

---
count: false

# Insurance Choice is Confusing

  - "More than half of all uninsured (56%) say they are likely to go to the marketplace to find a health plan in the future. But there are knowledge gaps."
  - "Six in 10 are confused about or have not heard of the tax credit, and more than 7 in 10 want one-on-one help understanding their insurance options"

.footnote[
  [Robert Wood Johnson Foundation, **Understanding the Uninsured Now**, 2015](https://www.rwjf.org/en/library/research/2015/06/understanding-the-uninsured-now.html)
]

---
# Choice of plans has big consequences