### 2021-03-25 finals/lunch-and-learn/lunch-learn-202103.Rmd
update lunch and learn presentation

---
title: "Insurance choice, decision assistance, and insurer steering"
subtitle: "Ian McCarthy & Evan Saltzman"
author: Emory University, Lunch and Learn
date: March 29, 2021 #"`r format(Sys.time(), '%d %B %Y')`"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
#knit: pagedown::chrome_print
---

<!-- Adjust some CSS code for font size, maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code, .remark-inline-code { 
    font-size: 20px;
}
</style>

<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(dplyr.summarise.inform = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T#, echo=F, warning=F, message=F
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, xaringanExtra)
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```

# Some decisions are hard

.center[
<iframe width="800" height="500" src="https://www.youtube.com/embed/oWXNf3JVWpI?start=146&end=271" frameborder="0" allowfullscreen></iframe>
]

---
# In the health insurance setting...

1. Insurance choice is confusing, particularly on the ACA exchanges

2. Size of the market introduces decision problems on a massive scale

---
# 1. Insurance choice is confusing

.center[
  ![](slate-2017.png)
]

.footnote[
  [Slate, 2017](https://slate.com/business/2017/08/americans-are-overpaying-for-insurance-because-obamacare-is-confusing.html)
]

---
count: false

# 1. Insurance choice is confusing
.center[
  ![](health-affairs-2017b.png)
]

.footnote[
  [Health Affairs, January, 2017](https://www.healthaffairs.org/doi/10.1377/hlthaff.2016.0472)
]

---
count: false

# 1. Insurance choice is confusing

.center[
  ![](health-affairs-2017.png)
]

.footnote[
  [Health Affairs, August, 2017](https://www.healthaffairs.org/doi/full/10.1377/hlthaff.2017.0085)
]

---
count: false

# 1. Insurance choice is confusing

- "More than half of all uninsured (56%) say they are likely to go to the marketplace to find a health plan in the future. But there are knowledge gaps."

- "Six in 10 are confused about or have not heard of the tax credit, and more than 7 in 10 want one-on-one help understanding their insurance options"

.footnote[
  [Robert Wood Johnson Foundation, **Understanding the Uninsured Now**, 2015](https://www.rwjf.org/en/library/research/2015/06/understanding-the-uninsured-now.html)
]

---
# 2. Potential scale

- 12.7 million enrollees on exchanges in 2016

- Down to 11.4 million in 2020

- Over 27 million remain uninsured

---
count: false

# 2. Potential scale

.pull-left[
Exchanges relatively large, but...

- Large number of people still uninsured

- Many eligible for some type of subsidy on exchanges
]

.pull-right[
  ![:scale 150%](kff-uninsured.png)
]

.footnote[
  [Kaiser Family Foundation, November, 2020](https://www.kff.org/policy-watch/millions-of-uninsured-americans-are-eligible-for-free-aca-health-insurance/)
]

---
# This paper

Lots of people participating in a very complicated decision. With this in mind, we ask the following questions:

1. Does some form of decision assistance "help" in plan selection?

2. Do insurers use brokers/agents to steer people to different plans?

3. What are the welfare effects of any such steering?

---
class: inverse

# Outline of talk

1. [Background on the ACA](#background)
2. [Data](#data)
3. [Dominated choices](#dominated)
4. [Effects of decision assistance](#decision-assist)
5. [Insurer steering](#steering)

---
class: inverse, center, middle
name: background

# Background on the ACA exchanges

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
class: clear, center, middle

![:scale 80%](https://media.giphy.com/media/Awb1k8lX8a3Re/giphy.gif)

---
# What are the exchanges?

- Online marketplace to purchase private health insurance for those without insurance through their employers or other public programs

- Intended to help fill the "coverage gap" for the 40+ million uninsured individuals in the U.S. (pre ACA)

- Complicated program with various premium and cost-sharing subsidies, specific to a household

---
# Metal tiers

.pull-left[
To help with this complex decision, plans organized around tiers based on actuarial value
]

--

.pull.right[
  ![:scale 50%](metal_tiers.jpg)
]

---
# Rating areas

.pull-left[
Insurers limited in ability to adjust premiums

- Smoking
- Age
- Rating area
]

--

.pull-right[
  ![:scale 100%](CA_Rating_Regions2.jpg)
]

---
# Popularity of the exchanges

.center[
![:scale 85%](aca-enrollment-graph.png)
]

---
class: inverse, center, middle
name: data

# Data

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Sources

Main enrollment come from Covered California under FOIA. This includes:
- Individual demographics and income
- Plan purchases
- Information on type of assistance (if any)

Supplemental data include:
- ACS and SIPP for predicting uninsured
- ACA program data (contribution percentages, rating areas, etc.)

---
# Final data

All analysis done at the household level...

- **Initial data:** 16.9 million households from 2014-2019
- Remove households with missing or inconsistent data
  - 14.6 million remaining
- Remove households predicted to be insured elsewhere
  - 8.3 million remaining
  - 4.5 million of which are new enrollees

---
# Enrollees over time

.center[
  ![:scale 50%](../../figures/enrollee_count.png)
]

---
# Insurers

.pull-left[
  ![](../../figures/insurer_stack_any.png)
]

.pull-right[
  ![](../../figures/insurer_stack_all.png)
]

---
# Metal tier

.pull-left[
  ![](../../figures/metal_stack_any.png)
]

.pull-right[
  ![](../../figures/metal_stack_all.png)

]

---
class: inverse, center, middle
name: dominated

# Analysis of dominated choices

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# What is a dominated choice?

<center>
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr"><a href="https://twitter.com/DaveEvansPhD/status/1234506064470716416">March 2, 2020</a></blockquote><script async src="//platform.twitter.com/widgets.js" charset="utf-8"></script>
</center>

---
count: false

# What is a dominated choice?

In our context...

- Subsidies available for households with incomes below 250% of the FPL
- At 150% of the FPL, larger subsidies are such that actuarial value of silver plans exceeds that of gold and platinum plans
- Gold and platinum plans dominated by silver plans for incomes $\leq$ 150% FPL

--

About 3.2% of households make a dominated choice in our data

---
# Analysis of dominated choices

.center[
  ![](../../figures/dominated_choice_regression.png)
]

---
class: inverse, center, middle
name: decision-assist

# Effect of decision assistance

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Consumer choice model

Household i's Utility for Plan j<br>

$\begin{eqnarray}U_{ij} &=& \alpha_i \underbrace{p_{ij}}_{cons.\\prem.} \hspace{0.075in} + \hspace{0.075in} \underbrace{x'_{j}}_{obs.\\prod.} \beta \hspace{0.075in} + \hspace{0.075in} \underbrace{d'_{i}}_{dem.\\char.} \varphi \hspace{0.075in} + \hspace{0.075in} \underbrace{\xi_{j}}_{unobs.\\prod.} \hspace{0.075in} + \hspace{0.075in} \underbrace{\epsilon_{ij}}_{error} \nonumber\end{eqnarray}$

where<br>

$\begin{eqnarray*}\alpha_i &=& \alpha + d'_i \gamma \end{eqnarray*}$

$\begin{eqnarray*}\underbrace{p_{ij}}_{cons.\\prem.} &=& \underbrace{r_{i}}_{rating\\ fact.} \underbrace{p_j}_{full\\prem.} \hspace{0.075in} - \hspace{0.075in} \underbrace{\max\{p_b - \mathrm{cap}_i,0\}}_{subsidy} \end{eqnarray*}$

---
count: false

# Consumer choice model

Utility for Outside Option (Uninsured)<br>

$\begin{eqnarray*} U_{i0} &=& \alpha_i \underbrace{\rho_i}_{\textrm{penalty}} \hspace{0.075in} + \hspace{0.075in}  \underbrace{\epsilon_{i0}}_{error} \end{eqnarray*}$

---
# Estimand of interest

- Want to identify the Average Treatment Effect on the Treated (ATT) of decision assistance

- Denote by $Pr[Y_{ij}^{1}]$ the probability of household $i$ selecting plan $j$ with decision assistance

- Denote by $Pr[Y_{ij}^{0}]$ the probability of household $i$ selecting plan $j$ without decision assistance

--

$$ATT = Pr[Y_{ij}^{1} | \text{Assist}] - Pr[Y_{ij}^{0} | \text{Assist}]$$

---
# Estimation

1. Estimate nested logit model among those without any decision assistance, $\hat{\theta}$

2. Use $\hat{\theta}$ to form predicted counterfactual, $\widehat{Pr[Y_{ij}^{0} | \text{Assist}]}$

3. Use predicted value in ATT estimate: $$\underbrace{Pr[Y_{ij}^{1} | \text{Assist}]}_{Observed} - \underbrace{\widehat{Pr[Y_{ij}^{0} | \text{Assist}]}}_{Predicted}$$

4. Aggregate to level of insurer and metal

---
count: false

# Estimation

- Focus only on new enrollees to avoid presence of inertia

- Estimate separately by year and rating area

---
# Inference

- Clustered bootstrap, by year and rating area

- 200 bootstrap replications

- Confidence interval based on the 5th and 95th percentile of bootstrap replications

---
# Effects of decision assistance

.pull-left[
  ![](../../figures/choice_metals.png)
]

.pull-right[
  ![](../../figures/choice_insurer.png)
]

---
class: inverse, center, middle
name: steering

# Analysis of insurer steering

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>